### 2014-08-15 response_to_reviewers.txt
Add response to reviewers

To all reviewers:
The manuscript has been drastically revised and polished. In particular, the
order of presentation and the motivation for the work has been expanded upon
greatly. Additionally, the example has been expanded upon significantly and the
steps shown more explicitly.

The electronic supplementary material should now be available and should be
executable by installing Python and SymPy (both free and open source, no other
dependencies). The simplest way to do this is to install one of the many
free scientific Python distributions which include both of these (and a number
of other useful tools). We recommend any of the following:
1) Enthought's Canopy: https://www.enthought.com/products/canopy/
2) Python XY: https://code.google.com/p/pythonxy/
3) Anacond Python: https://store.continuum.io/cshop/anaconda/
All are open source, well documented, and have free versions (and paid ones as
well with support), and have a userbase in the 10's-100's of thousands so
finding help online is generally quite easy.

To reviewer #1:
The new organization of the paper should address the concern you raised about
the presentation of linearized equations of motion without any deeper
motivation or introduction to the terms used.

Regarding the use of both independent states and dependent states, we address
this more in the introduction, the example, and the discussion of the revised
manuscript. Briefly, we formulate the equations which include both independent
and dependent states to allow for the possibility that which states are chosen
to be dependent/independent can be changed without having to redo the whole
symbolic formulation.

To reviewer #3:
Thank you for the references. We have since read all of them and highlighted in
the introduction the distinction between those works and our own.
Interestingly, we contacted Arend Schwab during the preparation of our
manuscript regarding references to linearization of constrained systems; his
response helpful, but did not include his own 2003 paper which you reference.
That aside, none of references you list employ Kane's method, nor do any
include discussion of allowing for non-trivial generalized speeds to be
selected (i.e., something other than d/dt(q_i) = u_i), nor are any of the
software packages to which the various authors refer open source -- with the
exception of the work by Minaker. The work by Minaker uses Lagrange's method,
and hence we feel is significantly distinct from our own; further he gives no
reference to the software which he refers to in his paper (Section 6),
presumably it is the software "EoM": http://137.207.14.230/vdc/index.html, but
we cannot be sure. If this is the software, it has no mailing list, uses no
source version control with a history of commits and contributions, and has no
automated test suite to verify it's correctness; it is essentially the project
of a single lab group and has no mechanism for participation by a wider
community. In contrast, the software we use to derive the equations of motion,
SymPy (sympy.org), is open source, has an extremely active mailing with
contributors from all over the world, and has been funded (through paying
students a small summer stipend) by Google for the last 6 years under the
Google Summer of Code program; it also has automated software tests for unit
tests a number example systems including the rolling disk.

Regarding the choice of independent speeds, the motivation for allowing the
choice of dependent speeds to be changed in the linearization procedure (even
though the dynamic equations, as you astutely observed, must be formed with
respect to a particular choice of dependent speeds) is so that the choice of
dependent speeds can be changed at the time of numerical evaluation rather than
at the time of symbolic derivation. We expand upon this more in the last
paragraph of the discussion, as well as provide an example of when it is useful
(computation of Lyapunov characteristic exponents for a system which a single
choice of dependent coordinates/speeds may be infeasible). Essentially, rather
than symbolically outputting the constrained equations of motion for every
possible choice of dependent states, it is possible to only output the
unconstrained equations of motion and the constraint Jacobian matrices, and
compute the "best" choice of dependent states (at runtime) and then
appropriately form the constrained dynamic equations (again, at runtime).