### 2014-08-15 response_to_reviewers_rev2.txt

Response to Reviewer #1 comments:

1. Change in revised abstract.

2. Changed in revised abstract to read: "(i.e. no DAE’s, dependent variables
are not substituted away)". We simply mean that algebraic elimination has not
been used to eliminate dependent variables.

3. We have removed the word "familiar".

4. We erased the duplicated word "system". The constraint equations are
involved as the reviewer mentions.

5. We replaced the word "types" with "levels".  Clarified
holonomic/nonholonomic and the use of time differentiated constraints at the
velocity and acceleration level.

6. We have clarified the text here.  Lagrange's method can be used for
nonoholonomic systems.  See Nonholonomic Mechanics and Control by Anthony Bloch
for extensive discussion of Lagrange's method applied to nonholonomic systems.
The two approaches (Kane's vs. Lagrange) are comparable but Kane's method
offers more flexibility on the choice of generalized speeds (i.e., u_i != dq_i
/ dt). The quoted sentence has been rewritten for clarity.

7. We have changed this equation for clarity.

8. Quoted text rewritten for clarity.  n and o are not required to be equal in
our approach, though this is commonly the case for simple holonomic systems.

9. ...

10. We have added a figure.

11.  The text has been clarified.

12.  We have revised this section to fix the reviewers comments and make the
section more clear. We do not consider the implementation/performance of a
symbolic matrix inversion.  The Jacobian of the right hand side of (44) can be
performed symbolically or numerically.  In this example, we obtained (44) by
symbolically separating and inverting terms in (43).  We symbolically obtained
(45) by computing the Jacobian using SymPy (www.sympy.org). The "steady rolling
condition" is exactly what it sounds like: the disk is rolling at a constant
spin rate, either in a straight line or along a circular path of constant
radius. See reference [14] by Oliver O'Reilly for a more detailed treatment of
rolling disks.

13. ...

14. ...

Response to Reviewer #3

1. The notation has been fixed in these equations.

2. The input description was clarified in the text following the table.

3. We have clarified this in the revised manuscript in the second paragraph
after Table 1.

4. Figure added.

5. Unnecessary step removed.

6. We have improved the discussion in the text. Having any incorrect
eigenvalues will result in an incorrect linear system response.  We do not
explore these incorrect results due to space limitations.

7. Thank you for catching this. The equation has been corrected.

8. This discrepancy has been clarified.

9. Thank you for catching this.  We added a definition for matrix M.