### 2014-08-16 response_to_reviewers_aug_2014.txt

Response to Editor.

Dear Editor(s),
We feel that in this revision, as in the previous revisions, we have made an
effort to address as many of Reviewer #1's comments as possible. We have valued
a number of the comments provided by Reviewer #1, agreeing that the suggested
changes help to clarify the motivation and methods in the paper.

Unfortunately, there are 2 comments from Reviewer #1 we do not feel are
reasonable to address, comments 7 and 12.

Comment 7 simply states that the reviewer finds our presentation frustrating,
without offering specific faults, mistakes, unclear sections, areas that could
use improvement, etc.  Additionally, some of the other comments (comment 3) are
the reason we have included such a thorough description of Kane's method. There
seems to be a breakdown where the reviewer is not understanding how we are
applying this method, and it is not clear on our end which part of Section 2
needs to be improved to address the reviewer's difficulties.

Comment 12 states that the reviewer found the parts of the manuscript that they
understood confusing, and then admits to being unable to follow the derivations
we have presented. While we cannot simplify the math in the derivations, one
can always attempt to provide more clarification, but only if the reviewer
states which areas they had trouble with or where throughout the derivation
they became lost. This comment as given to us, does not allow us to do
anything.

In summary, we have addressed as many of the comments as possible, but feel
that the other comments are not reasonable, and suggest that Reviewer #1 has
not made a good faith effort to either comprehend the paper, or communicate
where they had difficulty comprehending the paper. As Reviewer #2 seems to
approve of the manuscript, we would ask that you overlook Reviewer #1's
concerns when making the decision to publish or reject this work.

Thank you,
Luke Peterson
Gilbert Gede
Mont Hubbard

Response to Reviewer #1.

1. Abstract. “...but is compatible with any method as long as only ordinary
    differential equations are required to describe the system, and independent
    and dependent states remain in the final equations (i.e. no DAE's,
    dependent variables are not substituted away)”. The sentence, and
    especially its end, is still not clear to me.

Author Response: We have changed the sentence for clarity. It now more clearly
specifies the types of differential equations which can be used.  It also
clarifies that it was developed for use with Kane's Method when dependent
quantities have not been algebraically eliminated.

2. 1. 36-37 (page 1, lines 36-37). “Design of linear state feedback controllers
    and first order stability analysis of equilibria are two of the most common
    uses of multibody system dynamics models”. I am not sure about these are
    the most common uses.

Author Response: We removed the words which suggested that these were the most
common uses, instead stating that they are just common uses.

3. 2. 21-19. The paragraph is still unclear to me. Kane’s equations are the
   projections of the initial (expressed in dependent velocities and as such
   constraint reaction-induced) dynamic equations into the directions of
   generalized speeds, which are by assumption independent for a constrained
   system, and with these projections the constraint reactions are excluded
   from evidence. If the projected dynamic equations are still expressed in the
   dependent velocities (not in the generalized speeds), the number of the
   projected dynamic equations is smaller than the number of the velocities. In
   this context I do not understand the remark that “The kinematical and
   dynamical differential equations can be solved for the time derivatives of
   the generalized coordinates and generalized speeds, ...” as well as most of
   the content of the paragraph. My reservations raised in the previous review
   were not clarified.

Author Response: We believe you have some confusion regarding the different
possible ways to apply Kane's method, which is why we Section 2 clarifies how
we use Kane's method. One possible way to apply the method is when a
non-minimal set of speeds are used to define the velocities of the system; in
this case, the velocity constraint matrix B (equation 11 in the paper) can be
reworked into matrix A (equation 21). This A matrix is used to reduce the
number of equations for generalized speeds by the number of constrained speeds.
This is also outlined by Kane in [Kane 1985], in equations 4.4.3 and 4.4.11,
which shows the reduction in size of the equations for the generalized speeds
and that _all_ the speeds (dependent and independent) are still present.

4. 2. 20-27. I am not totally satisfied with the changes/clarifications made by
    the Authors. For example, the remark “Velocity constraints often appear
    where there is rolling without slip or there are closed kinematic loops” is
    unclear (if correct at all). Velocity constraints, or, better, constraint
    equations at the velocity level, are either nonholonomic constraints (often
    consequent to the rolling without slip) or differentiated with respect to
    time holonomic constraint equations (one source of such constraints are
    closed kinematic loops).

Author Response: We have rewritten this entire paragraph to ensure that the
description of each level of constraints is consistent, and more clear. We have
reworked the sentence in question to more clearly state that things like
rolling without slip lead to nonholonomic constraints and closed kinematic
loops lead to time differentiated holonomic constraints, both of which are
velocity level constraints.

5. 5. 25-50. The velocity constraints are introduced in Eq. (8) in a general
    (nonlinear) form f_v (q,u,t) = 0 . It is then said that “By assumption (why
    “by assumption”?), the velocity constraints are linear with respect to u,
    ... which implies Bu + fvt(q,t) = 0”. Firstly, in general, the constraint
    matrix B is not constant, i.e.  B(q) or even B(q,t). Then, why the velocity
    constraints were not introduced from the beginning in this “linear” form?

Author Response: We have clarified that the B matrix is a function of the
coordinates and time, even though it is written as only a single symbol (a
format which we use throughout the paper).

6. 8. 15 and 831. The meaning of the “exogenous inputs” is not well
    motivated/explained.

Author Response: We believe that the sentence "The s exogenous inputs include
control system inputs and/or external disturbance forces or torques applied to
the system." explains that the exogenous inputs include forces/torques which
are not internal to the system (e.g. a reaction force), but instead applied
from outside the system (a motor driving a body in the system).

7. A general remark to Section 2. While Kane’s method is generally well
    understood, the Authors presentation of its formulation is rather
    “frustrating”.

8. 10. 32-34. One of the velocity constraint equations introduced in Eq. (30)
    should be a time differentiated form of the configuration constraint of Eq.
    (25). It seems that it is Eq. (30c).  If so, it should rather read 
    r u2 s2 + u6  = 0 (not r u2 s3 + u6 = 0). Can you explain this?

Author Response: While it could be written this way, what we have presented is
correct. The method you are suggesting requires ^N v ^P* to be expressed in N.
Then, the N_x and N_y components would be 2 of the motion constraints, and the
time differentiated configuration constraint will equal the N_z component. This
can be simply verified by time differentiating the configuration constraint and
utilizing equation 34 to elimante the q_dots, so only u's are present. In
comparing this to the projection of ^N v ^P* in the N_z direction, you will
find these are equivalent. We believe what we have presented is simpler.

9. 10. 36-38. It is written: “These equations are linear in the ui terms,
    nonlinear in the qi terms, involve geometric system parameters (but not
    mass or inertial parameters), and do not explicitly involve \dot{qi}
    terms”. The remark that constraint equations do not involve inertial
    parameters is a cliché. According to Eq. (11), \dot{qi} terms are neither
    involved.

Author Response: Although it might be cliché, we feel explicitness is valuable,
and we do not wish to remove a sentence which might help a reader avoid
confusion.

10. 10. 39-118. The paragraph is unintelligible. The “diabolical set of
    coordinates and speeds” is just a set of dependent coordinates and speeds,
    resulted in one configuration constraint of Eq. (25) and three velocity
    constraints of Eq. (30) (one holonomic constraint of Eq.  (25) and two
    nonholonomic constraints of Eq. (30a) and (30b)). There are then five
    independent coordinates and three independent speeds. By choosing the
    independent coordinates and speeds, which may in general not be subsets of
    qi and ui, respectively, the constraint equations (25) and (30) are
    satisfied by identity.

Author Response: We have changed "diabolical" to "purposefully nonminimal".
That should clairfy that the example exists to show the technique on a
tractable system with a nonminimal set of coordinates and speeds, and to
highlight the approach, rather than suggest that we should be using these
coordinates and speeds on the rolling disk example.

11. 11. 14. Is Eq. (31c) correct? (see Remark 8).

Author Response: Yes it is correct, our response to comment 8 addresses this.
The code in the electronic supplementary material can be used to verify this.

12. A general remark. The complexity of the followed parts of the manuscript is
    not appealing to me. In fact, I was unable to follow the derivations. I may
    suppose that the same difficulties will have most of the potential readers
    of the paper, not mentioning those who would like to follow the procedures.