### 2014-08-29 response_to_reviewers_aug_2014.txt
Implement Mont's feedback from 3rd round of revisions

Response to Editor.

Dear Editor(s),
We feel that in this revision, as in the previous revisions, we have made an
effort to address as many of Reviewer #1's comments as possible. We have valued
a number of the comments provided by Reviewer #1, agreeing that the suggested
changes help to clarify the motivation and methods in the paper.

Unfortunately, there are 2 comments from Reviewer #1 we do not feel are
reasonable to address, comments 7 and 12.

Comment 7 simply states that the reviewer finds our presentation frustrating,
without offering specific faults, mistakes, unclear sections, areas that could
use improvement, etc.  Additionally, some of the other comments (comment 3) are
the reason we have included such a thorough description of Kane's method. There
seems to be a breakdown where the reviewer is not understanding how we are
applying this method, and it is not clear on our end which part of Section 2
needs to be improved to address the reviewer's difficulties.

Comment 12 states that the reviewer is confused by the parts of the manuscript
that they understood, and then admits to being unable to follow the derivations
we have presented. While we cannot simplify the math in the derivations, one
can always attempt to provide more clarification, but only if the reviewer
states which areas they had trouble with or where in the derivation they became
lost. This comment as given to us, does not allow us to do anything.

We would also like to add that many of Reviewer #1's comments seem to indicate
a shortage of effort to understand the paper. As an example, one of the authors
has advised a masters student in a summer project to implement this method in a
symbolic computer algebra system and within 2 months the student had
implemented the entirety of the algorithm, complete with examples, software
tests to verify mathematical correctness, and documentation. This student was
able to understand the mathematics presented in the paper with very little
guidance; fundamentally there is nothing in the derivation that is more
complicated than a partial derivative and a few matrix multiplications.

In summary, we have addressed as many of the comments as possible, but feel
that the other comments not addressable, and suggest that Reviewer #1 has
not made enough of an effort to either comprehend the paper, or communicate
where they had difficulty comprehending it. As Reviewer #2 seems to
approve of the manuscript, we would ask that you overlook Reviewer #1's
concerns when making the decision to publish or reject this work.

Thank you,
Dale Peterson
Gilbert Gede
Mont Hubbard

Response to Reviewer #1.

1. Abstract. “...but is compatible with any method as long as only ordinary
    differential equations are required to describe the system, and independent
    and dependent states remain in the final equations (i.e. no DAE's,
    dependent variables are not substituted away)”. The sentence, and
    especially its end, is still not clear to me.

Author Response: We have changed the sentence for clarity. It now more clearly
specifies the types of differential equations which can be used.  It also
clarifies that it was developed for use with Kane's Method when dependent
quantities have not been algebraically eliminated.

2. 1. 36-37 (page 1, lines 36-37). “Design of linear state feedback controllers
    and first order stability analysis of equilibria are two of the most common
    uses of multibody system dynamics models”. I am not sure about these are
    the most common uses.

Author Response: We removed the words which suggested that these were the most
common uses, instead stating that they are just common uses.

3. 2. 1-19. The paragraph is still unclear to me. Kane’s equations are the
   projections of the initial (expressed in dependent velocities and as such
   constraint reaction-induced) dynamic equations into the directions of
   generalized speeds, which are by assumption independent for a constrained
   system, and with these projections the constraint reactions are excluded
   from evidence. If the projected dynamic equations are still expressed in the
   dependent velocities (not in the generalized speeds), the number of the
   projected dynamic equations is smaller than the number of the velocities. In
   this context I do not understand the remark that “The kinematical and
   dynamical differential equations can be solved for the time derivatives of
   the generalized coordinates and generalized speeds, ...” as well as most of
   the content of the paragraph. My reservations raised in the previous review
   were not clarified.

We do not use 'generalized coordinates' or 'generalized speeds' to imply that
either set is minimal. In other words 'generalized' only implies that the
coordinates/speeds are for the entire system and are not Cartesian or
spherical coordinates necessarily.  The number of Kane's dynamical equations in our
formulation is o - m, where o is the number of generalized speeds and m is the
number of velocity (nonholonomic and time differentiated holonomic)
constraints. These o - m dynamical equations will in general involve 1) all n
coordinates and their time derivatives and 2) all o speeds and their time
derivatives. So despite the fact that the Newton-Euler equations have been
projected in the direction of the nonholonomic partial velocities (what you
refer to as 'the direction of the generalized speeds'), there are still more
unknowns (o generalized speed time derivatives) than there are equations (o -
m).  This issue can be addressed by 1) algebraically eliminating dependent
speed time derivatives to obtain o-m equations in o-m unknowns, or 2)
augmenting the o-m dynamical equations with the m acceleration constraint
equations to obtain o equations in o unknowns. Either way is mathematically
equivalent but dependent upon the specific problem at hand one may be
preferable to the other.

We have revised manuscript to be more explicit about what we mean by solving
the kinematical and dynamical differential equations for the time derivatives
of the coordinates and time derivatives of the generalize speeds. For this
dynamical differential equation:

-m*du/dt + f = 0

then by "solve the dynamical differential equations for the time derivatives of
the generalized speeds" we mean:

du/dt = f / m

We welcome specific suggestions about how this can be made more clear.

Section 2 was introduced specifically to clarify how we use Kane's method since
some of our notation deviates from his presentation in his 1985 book. When the
system has velocity constraints (i.e., a non-minimal set of generalized speeds
are introducted, or the system has time differentiated holonomic constraints),
the velocity constraint matrix B (equation 11 in the paper) can manipulated to
form matrix A (equation 21 in the paper). This A matrix is used to reduce the
number of equations for generalized speeds by the number of constrained speeds.
This is also outlined by Kane in [Kane 1985], in equations 4.4.3 and 4.4.11,
which shows the reduction in size of the equations for the generalized speeds
and that _all_ the speeds (dependent and independent) are still present.

4. 2. 20-27. I am not totally satisfied with the changes/clarifications made by
    the Authors. For example, the remark “Velocity constraints often appear
    where there is rolling without slip or there are closed kinematic loops” is
    unclear (if correct at all). Velocity constraints, or, better, constraint
    equations at the velocity level, are either nonholonomic constraints (often
    consequent to the rolling without slip) or differentiated with respect to
    time holonomic constraint equations (one source of such constraints are
    closed kinematic loops).

Author Response: We have rewritten this entire paragraph to ensure that the
description of each level of constraints is consistent, and more clear. We have
reworked the sentence in question to more clearly state that things like
rolling without slip lead to nonholonomic constraints and closed kinematic
loops lead to time differentiated holonomic constraints, both of which are
velocity level constraints.

5. 5. 25-50. The velocity constraints are introduced in Eq. (8) in a general
    (nonlinear) form f_v (q,u,t) = 0 . It is then said that “By assumption (why
    “by assumption”?), the velocity constraints are linear with respect to u,
    ... which implies Bu + fvt(q,t) = 0”. Firstly, in general, the constraint
    matrix B is not constant, i.e.  B(q) or even B(q,t). Then, why the velocity
    constraints were not introduced from the beginning in this “linear” form?

Author Response: We have clarified that the B matrix is a function of the
coordinates and time, even though it is written as only a single symbol (a
format which we use throughout the paper).

6. 8. 15 and 831. The meaning of the “exogenous inputs” is not well
    motivated/explained.

Author Response: We believe that the sentence "The s exogenous inputs include
control system inputs and/or external disturbance forces or torques applied to
the system." explains that the exogenous inputs include forces/torques which
are not internal to the system (e.g. a reaction force), but instead applied
from outside the system (a motor driving a body in the system).

7. A general remark to Section 2. While Kane’s method is generally well
    understood, the Authors presentation of its formulation is rather
    “frustrating”.

8. 10. 32-34. One of the velocity constraint equations introduced in Eq. (30)
    should be a time differentiated form of the configuration constraint of Eq.
    (25). It seems that it is Eq. (30c).  If so, it should rather read
    r u2 s2 + u6  = 0 (not r u2 s3 + u6 = 0). Can you explain this?

Author Response: While it could be written this way, what we have presented is
correct. The method you are suggesting requires ^N v ^P* to be expressed in N.
Then, the N_x and N_y components would be 2 of the motion constraints, and the
time differentiated configuration constraint will equal the N_z component. This
can be simply verified by time differentiating the configuration constraint and
utilizing equation 34 to eliminate the q_dots, so only u's are present. In
comparing this to the projection of ^N v ^P* in the N_z direction, you will
find these are equivalent. We believe what we have presented is simpler.

9. 10. 36-38. It is written: “These equations are linear in the ui terms,
    nonlinear in the qi terms, involve geometric system parameters (but not
    mass or inertial parameters), and do not explicitly involve \dot{qi}
    terms”. The remark that constraint equations do not involve inertial
    parameters is a cliché. According to Eq. (11), \dot{qi} terms are neither
    involved.

Author Response: Although it might be a cliché, we feel explicitness is
valuable, and we do not wish to remove a sentence which might help a reader
avoid confusion.

10. 10. 39-118. The paragraph is unintelligible. The “diabolical set of
    coordinates and speeds” is just a set of dependent coordinates and speeds,
    resulted in one configuration constraint of Eq. (25) and three velocity
    constraints of Eq. (30) (one holonomic constraint of Eq.  (25) and two
    nonholonomic constraints of Eq. (30a) and (30b)). There are then five
    independent coordinates and three independent speeds. By choosing the
    independent coordinates and speeds, which may in general not be subsets of
    qi and ui, respectively, the constraint equations (25) and (30) are
    satisfied by identity.

Author Response: We have changed "diabolical" to "purposefully nonminimal".
That should clarfy that the example exists to show the technique on a
tractable system with a non-minimal set of coordinates and speeds, and to
highlight the approach, rather than suggest that we should be using these
coordinates and speeds on the rolling disk example.

11. 11. 14. Is Eq. (31c) correct? (see Remark 8).

Author Response: Yes it is correct, our response to comment 8 addresses this.
The code in the electronic supplementary material can be used to verify this.

12. A general remark. The complexity of the followed parts of the manuscript is
    not appealing to me. In fact, I was unable to follow the derivations. I may
    suppose that the same difficulties will have most of the potential readers
    of the paper, not mentioning those who would like to follow the procedures.