### 2014-08-29 response_to_editors_aug_2014.txt
Separate responses to review and editor

Response to Editor.

Dear Editor(s),
We feel that in this revision, as in the previous revisions, we have made an
effort to address as many of Reviewer #1's comments as possible. We have valued
a number of the comments provided by Reviewer #1, agreeing that the suggested
changes help to clarify the motivation and methods in the paper.

Unfortunately, there are 2 comments from Reviewer #1 we do not feel are
reasonable to address, comments 7 and 12.

Comment 7 simply states that the reviewer finds our presentation frustrating,
without offering specific faults, mistakes, unclear sections, areas that could
use improvement, etc.  Additionally, some of the other comments (comment 3) are
the reason we have included such a thorough description of Kane's method. There
seems to be a breakdown where the reviewer is not understanding how we are
applying this method, and it is not clear on our end which part of Section 2
needs to be improved to address the reviewer's difficulties.

Comment 12 states that the reviewer is confused by the parts of the manuscript
that they understood, and then admits to being unable to follow the derivations
we have presented. While we cannot simplify the math in the derivations, one
can always attempt to provide more clarification, but only if the reviewer
states which areas they had trouble with or where in the derivation they became
lost. This comment as given to us, does not allow us to do anything.

We would also like to add that many of Reviewer #1's comments seem to indicate
a shortage of effort to understand the paper. As an example, one of the authors
has advised a masters student in a summer project to implement this method in a
symbolic computer algebra system and within 2 months the student had
implemented the entirety of the algorithm, complete with examples, software
tests to verify mathematical correctness, and documentation. This student was
able to understand the mathematics presented in the paper with very little
guidance; fundamentally there is nothing in the derivation that is more
complicated than a partial derivative and a few matrix multiplications.

In summary, we have addressed as many of the comments as possible, but feel
that the other comments not addressable, and suggest that Reviewer #1 has
not made enough of an effort to either comprehend the paper, or communicate
where they had difficulty comprehending it. As Reviewer #2 seems to
approve of the manuscript, we would ask that you overlook Reviewer #1's
concerns when making the decision to publish or reject this work.

Thank you,
Dale Peterson
Gilbert Gede
Mont Hubbard